canaccessR: An open data product for analyzing transportation accessibility to employment and grocery stores in Canada's largest metropolitan areas.

### 2025-03-11 paper/Manuscript-Data-Package.Rmd
create plot and finish draft
Create travel time map figure; include figure and brief description at the first version of the draft; finish first version of the draft.

keywords: Public transit accessibility; open data products (ODPs); R data package; travel time matrices.
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: 72
---

@@

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. 
The Table @tbl-table_1 below summarizes the travel time estimates.
The travel time datasets contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DA's in the sample and 18,519,897 pairs from population to groceries stores.
Considering all areas combined, the travel times estimated to jobs have an average of 78, standard deviation of 25, 1st quantile of 59, median of 80, and 3rd quantile of 99.
For travel times to groceries stores, these same statistics are the 76, 26, 56, 76, and 97, respectively.

```{r}
#| label: tbl-table_1
#| echo: false
#| message: false

@@

This section exemplifies how the package can be used through visual representation.

>>>>

keywords: Public transit accessibility; open data products (ODPs); R data package; travel time matrices.
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
    fig_caption: yes
header-includes: \usepackage{float}
editor_options: 
  markdown: 
    wrap: 72
---

@@

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. 
The Table @tbl-table_1 below summarizes the travel time estimates.
The travel time datasets contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DA's in the sample and 18,519,897 pairs from population to groceries stores.
Considering all areas combined, the mean travel time to jobs was 80 and 76 for groceries stores.

```{r, fig.pos="H"}
#| label: tbl-table_1
#| echo: false
#| message: false

@@

This section exemplifies how the package can be used through visual representation.
In Figure @fig-travel_time_emp_grc_plot, we present the spatial representation of the travel time matrices for the metropolitan region of Montréal.
In it, we see the median travel time from each DA to employment (left) and to groceries stores (right).
The plot shows that moving away from the city core increases the necessary travel time by public transit to reach employment locations and groceries stores.

```{r, fig.pos="H"}
#| label: fig-travel_time_emp_grc_plot
#| fig-cap: "Estimated median travel time (minutes) per Dissemination Area to jobs (left) and groceries stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021). DA and planning boundaries of the Montréal metropolitan area (Statistics Canada, 2016)."
#| out-width: 100%
#| echo: false
knitr::include_graphics("../figures/patch_tt_emp_grc.jpg")
```

### 2025-03-11 paper/Manuscript-Data-Package.tex
create plot and finish draft
Create travel time map figure; include figure and brief description at the first version of the draft; finish first version of the draft.

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}

@@

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. The Table \citet{tbl-table_1}
below summarizes the travel time estimates. The travel time datasets
contain, in total, 97,784,850 origin-destination pairs (observations)
from population to employment locations for all the DA's in the sample
and 18,519,897 pairs from population to groceries stores. Considering
all areas combined, the travel times estimated to jobs have an average
of 78, standard deviation of 25, 1st quantile of 59, median of 80, and
3rd quantile of 99. For travel times to groceries stores, these same
statistics are the 76, 26, 56, 76, and 97, respectively.

@@

This section presents some potential applications of the data package.

>>>>

\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}

@@

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. The Table \citet{tbl-table_1}
below summarizes the travel time estimates. The travel time datasets
contain, in total, 97,784,850 origin-destination pairs (observations)
from population to employment locations for all the DA's in the sample
and 18,519,897 pairs from population to groceries stores. Considering
all areas combined, the mean travel time to jobs was 80 and 76 for
groceries stores.

@@

This section exemplifies how the package can be used through visual
representation. In Figure \citet{fig-travel_time_emp_grc_plot}, we
present the spatial representation of the travel time matrices for the
metropolitan region of Montréal. In it, we see the median travel time
from each DA to employment (left) and to groceries stores (right). The
plot shows that moving away from the city core increases the necessary
travel time by public transit to reach employment locations and
groceries stores.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{../figures/patch_tt_emp_grc} \caption{Estimated median travel time (minutes) per Dissemination Area to jobs (left) and groceries stores (right). Public transit travel times are calculated using {r5r} (Pereira et al., 2021). DA and planning boundaries of the Montréal metropolitan area (Statistics Canada, 2016).}\label{fig:fig-travel_time_emp_grc_plot}
\end{figure}

A more thorough example of the package's use can be found on the School
of Cities' recent report on Canada's Urban Infrastructure Deficit. In
its 11th Chapter, we use the travel time matrices to estimate
accessibility metrics to jobs and groceries stores before and after the
pandemic \citep{pargaDemocraticAccessOur2024}. We then compare how
changes affected groups differently according to their spatial
distribution and income level, thus making explicit the connection of
the package's information and matters of equity in transportation. The
report is freely available for download at the State of Cities Summit
\href{https://stateofcitiessummit.ca/report}{website}.
